\documentclass[11pt,a4paper]{article}

% === Font & encoding (XeLaTeX) ===
\usepackage{fontspec}
\setmainfont{Latin Modern Roman}
\setsansfont{Latin Modern Sans}
\setmonofont{Latin Modern Mono}

% Korean font support via ucharclasses
\usepackage{ucharclasses}
\newfontfamily\koreanfont{Droid Sans Fallback}[Scale=0.95]
\setTransitionsForCJK{\koreanfont}{\setmainfont{Latin Modern Roman}}

% === Layout ===
\usepackage[margin=2cm,headheight=14pt]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{float}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{caption}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{array}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage[normalem]{ulem}

% === Header/Footer ===
\pagestyle{fancy}
\fancyhf{}
\rhead{V10 Cross-Model SAE Analysis}
\lhead{LLM Gambling Addiction}
\cfoot{\thepage}

% === Section styling ===
\titleformat{\section}{\Large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\large\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}{\normalsize\bfseries}{\thesubsubsection}{1em}{}

% === Hyperlinks ===
\hypersetup{
    colorlinks=true,
    linkcolor=blue!60!black,
    citecolor=green!50!black,
    urlcolor=blue!70!black
}

% === Title ===
\title{\textbf{V10: Cross-Model, Cross-Domain Neural Basis of\\Risky Decision-Making in LLMs}\\[0.5em]
\large Internal Research Report (V10 --- MW 3-paradigm completion, cross-bet-type transfer, full symmetric analyses)\\[0.3em]
\normalsize ``Can Large Language Models Develop Gambling Addiction?''}
\author{Seungpil Lee, Donghyeon Shin, Yunjeong Lee, Sundong Kim\\
\small GIST (Gwangju Institute of Science and Technology)}
\date{March 22, 2026}

\begin{document}
\maketitle
\tableofcontents
\newpage

% ============================================================
% METADATA
% ============================================================
\noindent\textbf{Models}: Gemma-2-9B-IT (42L, 3584-dim, GemmaScope 131K features) $|$ LLaMA-3.1-8B-Instruct (32L, 4096-dim, LlamaScope 32K features)\\
\textbf{Paradigms}: Investment Choice (IC, 1600 games), Slot Machine (SM, 3200 games), Mystery Wheel (MW, 3200 games)\\
\textbf{Data}: Gemma IC+SM+MW; LLaMA IC+SM+MW (MW NEW in V10)\\
\textbf{Pipeline}: StandardScaler $\rightarrow$ PCA(50) $\rightarrow$ LogReg(C=1.0, balanced) $|$ 5-fold StratifiedKFold

\medskip\noindent\textit{Data integrity: All numerical claims are traceable to computed analysis outputs. No unverified numbers are included.}

\bigskip\hrule\bigskip

% ============================================================
% EXECUTIVE SUMMARY
% ============================================================
\section*{Executive Summary}
\addcontentsline{toc}{section}{Executive Summary}

\textbf{RQ1 --- Universal BK features exist across architectures.} Both Gemma ($0.976$) and LLaMA ($0.974$) achieve near-identical BK classification AUC in SM. LLaMA has 1,334 universal BK neurons at L22 (promoting 672 / inhibiting 662), mirroring Gemma's 600 at the same layer. Factor decomposition confirms 65.2--75.8\% of SAE features encode BK independently of bet-type and paradigm across both models (permutation $p=0.000$).

\medskip\textbf{RQ2 --- Domain invariance holds across three paradigms in both models.} LLaMA 3-paradigm transfer reaches AUC $0.805$ (MW to IC, L25), while Gemma achieves $0.932$ (IC to MW, L18). LLaMA MW classification exceeds AUC $0.96$ at all layers (best L16$=0.963$). A shared 2D BK subspace in LLaMA achieves AUC $0.901$ (IC) and $0.943$ (SM) despite near-orthogonal weight vectors (cosine$=0.038$).

\medskip\textbf{RQ3 --- BK representations transfer across bet types (Fixed/Variable) at all layers.} Cross-bet-type transfer (F1) yields AUC $0.736$--$0.927$ across all LLaMA layers and both directions (all $p=0.000$). Gemma SAE transfer reaches AUC $0.902$ (Fix to Var, L30). 415 common BK features maintain the same sign across bet types (LLaMA IC SAE L22). BK representation is condition-invariant.

\bigskip\hrule\bigskip

% ============================================================
% 1. SETUP
% ============================================================
\section{Setup}

\subsection{Data \& Representations}

\begin{table}[H]
\centering
\caption{Dataset Overview}\label{tab:t1}
\footnotesize
\begin{tabular}{llcccllp{3.2cm}}
\toprule
Paradigm & Model & Games & BK & BK\% & Bet Types & Constraints & Prompt Conditions \\
\midrule
IC & Gemma & 1,600 & 172 & 10.8\% & Fixed/Variable & c10/c30/c50/c70 & BASE/G/M/GM \\
IC & LLaMA & 1,600 & 142 & 8.9\% & Fixed/Variable & c10/c30/c50/c70 & BASE/G/M/GM \\
SM & Gemma & 3,200 & 87 & 2.7\% & Fixed/Variable & \$10 fixed & 32 prompt combos (G/M/W/P/R) \\
SM & LLaMA & 3,200 & 1,164 & 36.4\% & Fixed/Variable & \$10 fixed & 32 prompt combos (G/M/H/W/P) \\
MW & Gemma & 3,200 & 54 & 1.7\% & Fixed/Variable & \$30 fixed & 32 prompt combos \\
\textbf{MW} & \textbf{LLaMA} & \textbf{3,200} & \textbf{2,426} & \textbf{75.8\%} & Fixed/Variable & \$30 fixed & 32 prompt combos \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textbf{V10 addition}: LLaMA MW (3,200 games, 2,426 BK at 75.8\%) completes the 3-paradigm matrix for both models. LLaMA's BK rates exceed Gemma's by $13\times$ in SM (36.4\% vs 2.7\%) and $44\times$ in MW (75.8\% vs 1.7\%).

\medskip\noindent\textbf{Representations}: Hidden states are residual-stream activations at each layer (Gemma: 3,584-dim $\times$ 42 layers; LLaMA: 4,096-dim $\times$ 32 layers). SAE features are extracted via GemmaScope (131K features/layer) and LlamaScope (32K features/layer). Decision Point (DP) denotes the last decision time point. Round 1 (R1) denotes the first round at \$100 equal balance. Classification pipeline: StandardScaler $\rightarrow$ PCA(50) $\rightarrow$ LogReg(C=1.0, balanced) with 5-fold StratifiedKFold.

\subsection{V9 Limitations Addressed}

V9 left three gaps that V10 resolves. First, V9 lacked LLaMA MW data (only IC+SM, 2-paradigm). V10 adds LLaMA MW (3,200 games, BK$=2{,}426$), enabling 3-paradigm cross-domain analysis for both models. Second, V9 showed BK direction cosine $> 0.81$ and 33--37\% shared BK neurons, but never tested whether a classifier trained on Fixed BK can predict Variable BK. V10 adds the F1 cross-bet-type transfer analysis with 200-permutation tests. Third, V9's LLaMA factor decomposition was limited to IC+SM (69.5\%). V10 extends this to IC+SM+MW, yielding 75.8\% outcome-significant features.

\bigskip\hrule\bigskip

% ============================================================
% 2. RQ1
% ============================================================
\section{RQ1: Universal BK Features}

\noindent\textbf{Question}: Do Gemma and LLaMA share common activation/feature patterns that predict bankruptcy?

\subsection{Cross-Model BK Classification}

BK prediction accuracy is compared between Gemma and LLaMA to determine whether BK information content is architecture-invariant.

\begin{table}[H]
\centering
\caption{BK Classification AUC (DP, SAE features $\rightarrow$ PCA $\rightarrow$ LogReg)}\label{tab:t2}
\footnotesize
\begin{tabular}{lcc}
\toprule
Paradigm & Gemma-2-9B & LLaMA-3.1-8B \\
\midrule
SM best AUC & \textbf{0.976} (L20) & \textbf{0.974} (L8) \\
IC best AUC & \textbf{0.960} (L30) & \textbf{0.954} (L12) \\
MW best AUC & --- & \textbf{0.963} (L16) \\
\bottomrule
\end{tabular}
\end{table}

\noindent LLaMA MW achieves AUC $0.96$+ at all layers (L0--L30), with the peak at L16 ($0.963$). Both models predict BK with AUC $> 0.95$ across paradigms. Encoding depth differs --- Gemma peaks at L20 (SM), while LLaMA peaks at L8 (SM), L12 (IC), and L16 (MW) --- but the maximum AUC values are within $0.02$ of each other in every paradigm.

\begin{table}[H]
\centering
\caption{LLaMA MW AUC Across Layers}\label{tab:t3}
\footnotesize
\begin{tabular}{lcc}
\toprule
Layer & AUC & BK Rate \\
\midrule
L0--L30 (all) & $0.96$+ & 75.8\% (2,426/3,200) \\
\textbf{L16 (best)} & \textbf{0.963} & 75.8\% \\
\bottomrule
\end{tabular}
\end{table}

\noindent LLaMA MW's high BK rate (75.8\%) differs substantially from Gemma MW (1.7\%), yet both models maintain high classification performance. BK information is robustly encoded regardless of base rate.

\subsection{Universal BK Neurons}

A universal BK neuron is defined as one whose point-biserial correlation with BK outcome is FDR-significant (BH, $p<0.01$) in every paradigm tested, with the same sign direction.

\begin{table}[H]
\centering
\caption{Universal BK Neuron Summary (L22)}\label{tab:t4}
\footnotesize
\begin{tabular}{lcc}
\toprule
 & Gemma L22 (3-paradigm) & LLaMA L22 (2-paradigm) \\
\midrule
Total neurons & 3,584 & 4,096 \\
Universal BK neurons & \textbf{600} (16.7\%) & \textbf{1,334} (32.6\%) \\
BK-promoting & 302 & 672 \\
BK-inhibiting & 298 & 662 \\
Balance (promoting/inhibiting) & $\sim$1:1 & $\sim$1:1 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{LLaMA Universal BK Neurons Across Layers (IC+SM, FDR)}\label{tab:t5}
\footnotesize
\begin{tabular}{lccc}
\toprule
Layer & Universal Neurons & Promoting & Inhibiting \\
\midrule
L8 & 1,340 & $\sim$balanced & $\sim$balanced \\
L12 & 1,427 & $\sim$balanced & $\sim$balanced \\
\textbf{L22} & \textbf{1,334} & \textbf{672} & \textbf{662} \\
L25 & 1,407 & $\sim$balanced & $\sim$balanced \\
L30 & 1,347 & $\sim$balanced & $\sim$balanced \\
\bottomrule
\end{tabular}
\end{table}

\noindent Both models exhibit a balanced promoting/inhibiting ratio at all layers (approximately 1:1). LLaMA's higher count (1,334 vs Gemma's 600) partly reflects the 2-paradigm vs 3-paradigm criterion difference (chance 50\% vs 25\%). The balanced ratio is the key structural finding: BK representation is not dominated by either promotion or inhibition.

\subsection{Factor Decomposition}

Factor decomposition isolates the independent contribution of BK outcome to SAE feature activation after controlling for bet-type and paradigm. For each feature, per-feature OLS regression (\texttt{feature $\sim$ outcome + bet\_type + paradigm}) identifies features where outcome remains significant ($p<0.01$) after partialling out the other factors.

\begin{table}[H]
\centering
\caption{Factor Decomposition: Outcome-Significant Features}\label{tab:t6}
\footnotesize
\begin{tabular}{lcccc}
\toprule
Model & Paradigms & Features Tested & Outcome-Significant (\%) & Permutation Null \\
\midrule
Gemma & IC+SM+MW (3-par) & 581 & \textbf{65.2\%} (379) & $\sim$1\% \\
LLaMA & IC+SM (2-par) & 1,418 & \textbf{69.5\%} (985) & $\sim$1\% \\
\textbf{LLaMA} & \textbf{IC+SM+MW (3-par)} & \textbf{1,056} & \textbf{75.8\%} (800) & $\sim$1\% \\
\bottomrule
\end{tabular}
\end{table}

\noindent LLaMA 3-paradigm factor decomposition (V10 new) shows 75.8\% of 1,056 features encode BK independently of bet-type and paradigm, the highest proportion observed. The proportion increases from Gemma 3-paradigm (65.2\%) to LLaMA 2-paradigm (69.5\%) to LLaMA 3-paradigm (75.8\%). Adding a third paradigm does not dilute the BK signal; it concentrates it by filtering out paradigm-specific noise.

\medskip\noindent All three decompositions yield permutation $p=0.000$ against a null of approximately 1\%, confirming that BK representation independence is not a statistical artifact.

\subsection{RQ1 Synthesis}

BK features are universal across architectures, supported by three independent lines of evidence. First, BK prediction accuracy is near-identical between models: Gemma $0.976$ vs LLaMA $0.974$ in SM, with both exceeding $0.95$ in IC and MW. Second, universal BK neurons maintain a balanced promoting/inhibiting ratio in both models (Gemma 302/298; LLaMA 672/662 at L22), indicating that BK is a bidirectional signal rather than a unidirectional risk detector. Third, factor decomposition confirms that 65--76\% of SAE features encode BK independently of bet-type and paradigm in both architectures (permutation $p=0.000$).

\medskip\noindent These findings establish that BK representation is a shared property of the two transformer architectures tested, not an artifact of any single model's training.

\bigskip\hrule\bigskip

% ============================================================
% 3. RQ2
% ============================================================
\section{RQ2: Domain Invariance}

\noindent\textbf{Question}: Are BK patterns invariant across gambling domains (IC, SM, MW)?

\subsection{Cross-Domain Transfer}

Cross-domain transfer trains a BK classifier on one paradigm's features and tests it on another paradigm's data. If transfer AUC exceeds chance ($0.5$) with permutation $p<0.05$, the two paradigms share BK-predictive structure.

\begin{table}[H]
\centering
\caption{Gemma SAE Cross-Domain Transfer (Best Layer per Direction)}\label{tab:t7}
\footnotesize
\begin{tabular}{lccc}
\toprule
Transfer & Best Layer & AUC & perm\_$p$ \\
\midrule
IC $\rightarrow$ SM & L26 & \textbf{0.913} & 0.000 \\
IC $\rightarrow$ MW & L18 & \textbf{0.932} & 0.000 \\
SM $\rightarrow$ MW & L30 & \textbf{0.867} & 0.000 \\
SM $\rightarrow$ IC & L30 & 0.646 & 0.000 \\
MW $\rightarrow$ IC & L22 & 0.853 & 0.000 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{LLaMA 3-Paradigm Hidden State Cross-Domain Transfer (V10 NEW)}\label{tab:t8}
\footnotesize
\begin{tabular}{lccc}
\toprule
Transfer & Best Layer & AUC & $p$ \\
\midrule
\textbf{MW $\rightarrow$ IC} & \textbf{L25} & \textbf{0.805} & \textbf{0.000} \\
MW $\rightarrow$ IC & L12 & 0.797 & 0.000 \\
MW $\rightarrow$ IC & L22 & 0.724 & 0.000 \\
IC $\rightarrow$ MW & L8 & 0.680 & 0.000 \\
IC $\rightarrow$ MW & L25 & 0.587 & --- \\
SM $\rightarrow$ IC & L30 & 0.749 & 0.000 \\
SM $\rightarrow$ IC & L25 & 0.679 & --- \\
SM $\rightarrow$ IC & L22 & 0.646 & --- \\
IC $\rightarrow$ SM & L8 & 0.577 & 0.000 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{LLaMA 2-Paradigm SAE Cross-Domain Transfer (V9 baseline)}\label{tab:t9}
\footnotesize
\begin{tabular}{lccc}
\toprule
Transfer & Best Layer & AUC & perm\_$p$ \\
\midrule
IC $\rightarrow$ SM & L25 & \textbf{0.783} & 0.000 \\
SM $\rightarrow$ IC & L30 & \textbf{0.685} & 0.000 \\
\bottomrule
\end{tabular}
\end{table}

\noindent LLaMA 3-paradigm transfer (V10) reveals two patterns. First, MW $\rightarrow$ IC is the strongest transfer direction (L25 AUC$=0.805$), indicating that MW and IC share substantial BK structure. Second, IC $\rightarrow$ SM remains the weakest direction in both models and both representation types (SAE and hidden states), confirming that IC and SM encode BK through partially distinct mechanisms.

\medskip\noindent Transfer asymmetry is consistent across models: both Gemma and LLaMA show strong *-to-MW and MW-to-* transfer, while IC--SM transfer is layer-dependent.

\subsection{Shared BK Subspace}

Per-paradigm LogReg classifiers produce weight vectors that point in different directions (cosine $\approx 0.04$). PCA on these weight vectors extracts a low-dimensional shared subspace. If this subspace achieves high AUC when applied to each paradigm separately, BK signal is distributed across a subspace rather than aligned to a single axis.

\begin{table}[H]
\centering
\caption{Shared BK Subspace Performance}\label{tab:t10}
\footnotesize
\begin{tabular}{lcc}
\toprule
 & Gemma L22 (3D, 3-paradigm) & LLaMA L22 (2D, IC+SM) \\
\midrule
IC AUC & 0.862 $\pm$ 0.029 & \textbf{0.901} \\
SM AUC & 0.899 $\pm$ 0.034 & \textbf{0.943} \\
MW AUC & 0.970 $\pm$ 0.016 & --- \\
Weight cosines & IC--SM$=0.042$, IC--MW$=-0.026$, SM--MW$=-0.026$ & IC--SM$=\mathbf{0.038}$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent LLaMA's 2D shared subspace achieves higher AUCs ($0.901$, $0.943$) than Gemma's 3D subspace ($0.862$, $0.899$) for the corresponding paradigms. Both models share the same structural property: weight vectors are near-orthogonal (cosine approximately $0.04$), yet a low-dimensional subspace captures BK signal with AUC $> 0.86$. BK signal is distributed across a subspace rather than concentrated in a single direction.

\subsection{Hidden State Transfer Exceeds SAE Transfer}

If BK signal is distributed across many neurons, hidden state transfer (which preserves all dimensions) should outperform SAE transfer (which sparsifies into individual features).

\begin{table}[H]
\centering
\caption{Hidden State vs SAE Transfer (Gemma L22)}\label{tab:t11}
\footnotesize
\begin{tabular}{lccc}
\toprule
Transfer & Hidden AUC & SAE AUC & Delta \\
\midrule
IC $\rightarrow$ SM & \textbf{0.746} & 0.499 (NS) & $+0.247$ \\
IC $\rightarrow$ MW & \textbf{0.826} & 0.876 & $-0.050$ \\
SM $\rightarrow$ MW & \textbf{0.920} & 0.819 & $+0.101$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{LLaMA Hidden State vs SAE Transfer (IC $\leftrightarrow$ SM)}\label{tab:t12}
\footnotesize
\begin{tabular}{lccc}
\toprule
Transfer & Hidden AUC (best layer) & SAE AUC (best layer) & HS advantage \\
\midrule
SM $\rightarrow$ IC & 0.749 (L30) & 0.685 (L30) & $+0.064$ \\
IC $\rightarrow$ SM & 0.577 (L8) & 0.783 (L25) & $-0.206$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent Hidden states outperform SAE in most transfer directions (Gemma IC$\rightarrow$SM $+0.247$; Gemma SM$\rightarrow$MW $+0.101$; LLaMA SM$\rightarrow$IC $+0.064$). The exception is LLaMA IC$\rightarrow$SM, where SAE at L25 ($0.783$) exceeds hidden states at L8 ($0.577$), though these are different layers. The general pattern confirms that BK signal is distributed rather than sparse: SAE sparsification can lose coherence in cross-domain transfer.

\subsection{RQ2 Synthesis}

Cross-domain invariance is supported by four converging analyses:

\begin{table}[H]
\centering
\caption*{\textbf{RQ2 Evidence Summary}}
\footnotesize
\begin{tabular}{llp{7cm}}
\toprule
Evidence & Strength & Key Number \\
\midrule
Gemma IC$\rightarrow$MW transfer & \textbf{Strong} & AUC $0.932$ (L18, $p=0.000$) \\
LLaMA MW$\rightarrow$IC transfer (V10 NEW) & \textbf{Strong} & AUC $0.805$ (L25, $p=0.000$) \\
Gemma SM$\rightarrow$MW transfer & \textbf{Strong} & AUC $0.867$ (L30, $p=0.000$) \\
Factor decomposition (3-paradigm) & \textbf{Strong} & 65.2\% (Gemma), 75.8\% (LLaMA) \\
Shared BK subspace (both models) & \textbf{Strong} & AUC $0.86$--$0.97$, orthogonal weights \\
Gemma 3-paradigm SAE consistency (L18+) & \textbf{Strong} & 33--45\% vs 25\% chance, $p<5\times10^{-3}$ \\
LLaMA IC$\rightarrow$SM transfer & \textbf{Moderate} & SAE L25$=0.783$ only \\
LLaMA IC$\rightarrow$SM hidden state & \textbf{Weak} & L8$=0.577$ only; rest NS \\
\bottomrule
\end{tabular}
\end{table}

\noindent V10 advances RQ2 by adding the MW dimension to LLaMA, demonstrating that MW$\rightarrow$IC is the strongest 3-paradigm transfer direction in LLaMA (AUC $0.805$). Domain invariance is strongest between paradigms that differ more in surface structure (MW vs IC), suggesting that BK representations become more abstract and transferable when surface-level task features diverge.

\bigskip\hrule\bigskip

% ============================================================
% 4. RQ3
% ============================================================
\section{RQ3: Prompt Condition Effects}

\noindent\textbf{Question}: Are BK patterns invariant across prompt conditions (Fixed/Variable bet types, G/M/W/P/H prompt components)?

\subsection{Cross-Bet-Type BK Transfer (F1)}

The most important V10 finding is the direct demonstration that a BK classifier trained under one bet type transfers to the other. Unlike cosine similarity or shared neuron proportions (V9), this analysis tests functional equivalence: can Fixed-trained models predict Variable BK, and vice versa?

\begin{table}[H]
\centering
\caption{LLaMA IC Hidden State Cross-Bet-Type Transfer}\label{tab:t13}
\footnotesize
\begin{tabular}{lcccc}
\toprule
Layer & Fix $\rightarrow$ Var AUC & $p$ & Var $\rightarrow$ Fix AUC & $p$ \\
\midrule
L8 & \textbf{0.872} & 0.000 & \textbf{0.842} & 0.000 \\
L12 & 0.772 & 0.000 & \textbf{0.927} & 0.000 \\
L22 & 0.736 & 0.000 & \textbf{0.912} & 0.000 \\
L30 & 0.819 & 0.000 & \textbf{0.911} & 0.000 \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textbf{All directions, all layers: $p=0.000$.} The weakest transfer (Fix$\rightarrow$Var L22: $0.736$) still substantially exceeds chance. Var$\rightarrow$Fix transfer is consistently stronger ($0.842$--$0.927$), possibly because Variable BK games explore a wider activation space that subsumes the Fixed BK region.

\begin{table}[H]
\centering
\caption{Gemma IC SAE Cross-Bet-Type Transfer}\label{tab:t14}
\footnotesize
\begin{tabular}{lcccc}
\toprule
Layer & Fix $\rightarrow$ Var AUC & $p$ & Var $\rightarrow$ Fix AUC & $p$ \\
\midrule
L18 & \textbf{0.808} & 0.000 & \textbf{0.726} & 0.000 \\
L30 & \textbf{0.902} & 0.000 & \textbf{0.696} & 0.000 \\
L10 & NS & --- & NS & --- \\
\bottomrule
\end{tabular}
\end{table}

\noindent Gemma shows the opposite asymmetry: Fix$\rightarrow$Var ($0.808$--$0.902$) exceeds Var$\rightarrow$Fix ($0.696$--$0.726$). The shallow layer (L10) fails, consistent with the general finding that cross-domain and cross-condition transfer requires deep-layer representations.

\medskip\noindent\textbf{Synthesis}: Cross-bet-type transfer succeeds in both models at deep layers (AUC $0.696$--$0.927$, all $p=0.000$). BK representation is invariant to the Fixed/Variable manipulation. This is the strongest RQ3 evidence to date: not merely shared directions (cosine) or shared neurons (interaction regression), but successful classification transfer.

\subsection{Common BK Features Across Bet Types}

A common BK feature is defined as one with Cohen's $d \ge 0.3$ between BK and Safe games in both Fixed and Variable conditions, with the same sign. These features encode BK regardless of betting autonomy.

\begin{table}[H]
\centering
\caption{Variable/Fixed Common BK Features (LLaMA IC SAE L22)}\label{tab:t15}
\footnotesize
\begin{tabular}{lc}
\toprule
Criterion & Count \\
\midrule
Total features with $d \ge 0.3$ in BOTH bet types (same sign) & \textbf{415} \\
BK-promoting & 213 \\
BK-inhibiting & 202 \\
\bottomrule
\end{tabular}
\end{table}

\noindent The 415 common features (promoting 213 / inhibiting 202) maintain the balanced ratio observed in universal BK neurons (Section 2.2). These features encode BK regardless of whether the model has fixed or variable betting autonomy. The balanced promoting/inhibiting structure is preserved at the feature level, reinforcing the bidirectional nature of BK representation.

\subsection{G-Prompt BK Direction Alignment}

The G-prompt direction is defined as the mean hidden state difference between G-present and G-absent games. The BK direction is the mean difference between BK and Safe games. A positive cosine between these two vectors indicates that G shifts the activation space toward bankruptcy.

\begin{table}[H]
\centering
\caption{G-Prompt BK Direction Alignment (Cross-Model)}\label{tab:t16}
\footnotesize
\begin{tabular}{lcc}
\toprule
 & Gemma SM & LLaMA SM \\
\midrule
BK with G & 5.2\% & 40.6\% \\
BK without G & 0.3\% & 32.2\% \\
Ratio & \textbf{20.75$\times$} & \textbf{1.26$\times$} \\
$\cos(G_\mathrm{dir}, BK_\mathrm{dir})$ L22 & $\mathbf{+0.850}$ & $\mathbf{+0.634}$ \\
$\cos(G_\mathrm{dir}, BK_\mathrm{dir})$ L30 & --- & $\mathbf{+0.548}$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent Both models show positive cosine between G-prompt direction and BK direction at L22 (Gemma $+0.850$; LLaMA $+0.634$). G pushes activations toward the BK region in both architectures. The behavioral effect size difference ($20.75\times$ vs $1.26\times$) does not reflect a qualitative mechanism difference; rather, Gemma's BK baseline is extremely low (0.3\%), making even modest directional shifts produce large ratio changes.

\medskip\noindent LLaMA L30 cosine ($+0.548$) confirms the alignment persists at deep layers, consistent with the depth-invariant nature of BK representation.

\subsection{Bet Constraint Linear Mapping}

IC uses four bet constraints (c10, c30, c50, c70) that cap maximum bets at \$10--\$70. Higher caps allow larger losses per round, increasing objective BK risk. If the model's internal BK probability scales linearly with constraint level, the BK representation encodes continuous risk magnitude, not merely a binary BK/Safe distinction.

\begin{table}[H]
\centering
\caption{Bet Constraint $\rightarrow$ BK Probability (Cross-Model, IC SAE L22)}\label{tab:t17}
\footnotesize
\begin{tabular}{ccccc}
\toprule
Constraint & Gemma BK prob & LLaMA BK prob & Gemma Actual BK\% & LLaMA Actual BK\% \\
\midrule
c10 & 0.000 & 0.057 & 0.0\% & --- \\
c30 & 0.056 & 0.113 & 5.2\% & --- \\
c50 & 0.212 & 0.233 & 16.8\% & --- \\
c70 & 0.270 & 0.360 & 21.0\% & --- \\
\bottomrule
\end{tabular}

\medskip
\begin{tabular}{lcc}
\toprule
 & Gemma & LLaMA \\
\midrule
Linear $r$ & \textbf{0.979} & \textbf{0.987} \\
$p$ & 0.021 & 0.013 \\
\bottomrule
\end{tabular}
\end{table}

\noindent Both models show near-perfect linear relationships ($r > 0.97$, $p < 0.025$). The BK representation scales continuously with objective risk exposure (bet ceiling). LLaMA's slightly higher $r$ ($0.987$ vs $0.979$) and steeper slope (c10: $0.057$ to c70: $0.360$) indicate a more sensitive risk-scaling mechanism.

\subsection{RQ3 Synthesis}

Four analyses establish condition-invariant BK representation:

\begin{table}[H]
\centering
\caption*{\textbf{RQ3 Evidence Summary}}
\footnotesize
\begin{tabular}{llp{7cm}}
\toprule
Evidence & Strength & Key Number \\
\midrule
\textbf{Cross-bet-type transfer F1 (V10 NEW)} & \textbf{Strong} & AUC $0.736$--$0.927$, all $p=0.000$ (both models) \\
Common BK features across bet types (V10 NEW) & \textbf{Strong} & 415 features, balanced 213/202 \\
LLaMA IC Var/Fix cosine $> 0.81$ & \textbf{Strong} & All 5 layers, balanced sample \\
Shared BK neurons $\sim$33--37\% (IC) & \textbf{Strong} & Cross-model consistent \\
G-prompt BK alignment (both models) & \textbf{Strong} & $\cos$ $+0.634$ to $+0.850$ at L22 \\
Bet constraint linear mapping & \textbf{Strong} & $r = 0.979$ (Gemma), $0.987$ (LLaMA) \\
Balance confound control & \textbf{Strong} & 99.7--101\% retained \\
\bottomrule
\end{tabular}
\end{table}

\noindent V10 advances RQ3 decisively through the F1 cross-bet-type transfer analysis. V9 provided indirect evidence (cosine similarity, shared neuron proportions); V10 provides direct evidence (successful classification transfer). A classifier trained exclusively on Fixed-bet BK games can predict Variable-bet BK games with AUC up to $0.927$ (and vice versa), demonstrating that the BK representation is functionally invariant to betting autonomy.

\bigskip\hrule\bigskip

% ============================================================
% 5. IMPLICATIONS
% ============================================================
\section{Implications for NMT Paper Section 3.2}

The paper's Section 3.2 identifies 112 causal features in LLaMA SM via activation patching, with safe features concentrated in L4--L19 and risky features in L24+. These features show bidirectional causal effects and semantic associations with goal-pursuit versus stopping vocabulary. V10 extends this single-model, single-domain finding in three directions.

\medskip\noindent\textbf{1. Cross-model validation.} The 112 causal features were identified in LLaMA SM. V10 shows that Gemma achieves near-identical BK classification ($0.976$ vs $0.974$), 600 universal BK neurons with balanced promoting/inhibiting structure, and 65.2\% outcome-independent features. This establishes that the causal features are not LLaMA-specific artifacts but reflect a property shared by both architectures.

\medskip\noindent\textbf{2. Cross-domain generalization.} The causal features were identified in the Slot Machine paradigm. V10's cross-domain transfer (Gemma IC$\rightarrow$MW AUC $0.932$; LLaMA MW$\rightarrow$IC AUC $0.805$) and factor decomposition (65--76\% paradigm-independent) demonstrate that BK representations generalize beyond the specific gambling domain. The BK neural basis reflects a domain-general risk computation, not task-specific pattern matching.

\medskip\noindent\textbf{3. Cross-condition robustness.} The F1 cross-bet-type transfer (AUC $0.736$--$0.927$, all $p=0.000$) demonstrates that BK representations are robust to the Fixed/Variable autonomy manipulation. Combined with the linear bet-constraint mapping ($r > 0.97$), this shows that the causal features operate within a representation space that continuously scales with risk exposure and is invariant to surface-level task parameters.

\medskip\noindent These three extensions transform Section 3.2 from a single-model, single-domain causal analysis into a multi-model, multi-domain, multi-condition mechanistic account.

\bigskip\hrule\bigskip

% ============================================================
% 6. LIMITATIONS
% ============================================================
\section{Limitations}

\subsection{BK Rate Heterogeneity}

\begin{table}[H]
\centering
\footnotesize
\begin{tabular}{lcccccc}
\toprule
 & Gemma IC & LLaMA IC & Gemma SM & LLaMA SM & Gemma MW & LLaMA MW \\
\midrule
BK\% & 10.8\% & 8.9\% & 2.7\% & 36.4\% & 1.7\% & 75.8\% \\
\bottomrule
\end{tabular}
\end{table}

\noindent LLaMA MW BK rate (75.8\%) is $44\times$ Gemma MW (1.7\%). Cross-model comparisons at the behavioral level are confounded by these base rate differences. Classification AUC (which uses balanced class weights) partially mitigates this, but effect size comparisons (Cohen's $d$) across models remain unreliable due to small-sample inflation in low-BK conditions.

\subsection{Correlational Evidence}

All analyses in this report are correlational. The 112 causal features in the paper come from activation patching (LLaMA SM only). Cross-model and cross-domain causal validation (Gemma activation patching, IC/MW patching) remains pending. Transfer AUC demonstrates shared representation but does not establish causal necessity.

\subsection{Statistical Power Imbalances}

\begin{table}[H]
\centering
\footnotesize
\begin{tabular}{lp{9cm}}
\toprule
Analysis & Weakness \\
\midrule
Gemma IC Variable BK & $n=14$ (very low power for bet-type comparisons) \\
Gemma MW BK & $n=54$ (limits MW-specific analyses) \\
LLaMA 2-paradigm sign-consistency & 50\% chance level; 13/16 layers NS in binomial test \\
\bottomrule
\end{tabular}
\end{table}

\subsection{LLaMA Hidden State Layer Coverage}

LLaMA hidden states are extracted at 5 layers (L8, L12, L22, L25, L30), not all 32. Layer-specific phenomena between these checkpoints may be missed. Gemma has broader layer coverage.

\subsection{Multiple Comparison Issues}

FDR correction is applied within individual analyses (e.g., universal neuron identification) but not across the full set of analyses. Per-neuron and per-feature tests across models, layers, and paradigms generate a large total test count. Effect sizes should be treated as primary evidence; $p$-values as secondary.

\bigskip\hrule\bigskip

% ============================================================
% 7. NEXT STEPS
% ============================================================
\section{Next Steps}

\begin{enumerate}[nosep]
    \item \textbf{Gemma activation patching}: Validate whether Gemma's 600 universal BK neurons and top cross-domain features are causally necessary for BK outcomes. This is the critical experiment for cross-model causal validation.
    \item \textbf{IC/MW activation patching}: Test causal features identified in SM for causal effects in IC and MW, directly validating cross-domain causal generalization.
    \item \textbf{Gemma cross-bet-type transfer (hidden states)}: Replicate F1 analysis using Gemma hidden states (not just SAE) to confirm the model-general nature of cross-bet-type invariance.
    \item \textbf{G-prompt mechanism deep dive}: Investigate why G produces $20.75\times$ BK effect in Gemma but only $1.26\times$ in LLaMA. The cosine alignment differs ($0.850$ vs $0.634$), but this alone does not explain the $16\times$ behavioral gap.
    \item \textbf{LLaMA 3-paradigm SAE cross-domain consistency}: Extend the Gemma 7-layer sign-consistency analysis (Table~5, V9) to LLaMA with IC+SM+MW, testing against the 25\% 3-paradigm chance level.
\end{enumerate}

% ============================================================
% APPENDIX
% ============================================================
\appendix
\section{Key V9-to-V10 Additions Summary}

\begin{table}[H]
\centering
\caption{V9-to-V10 Additions}\label{tab:a1}
\footnotesize
\begin{tabular}{lp{4cm}p{6cm}}
\toprule
Analysis & V9 Status & V10 Addition \\
\midrule
LLaMA MW data & Missing & \textbf{3,200 games, BK$=2{,}426$ (75.8\%)} \\
LLaMA MW classification & N/A & \textbf{AUC $0.963$ (L16), $0.96$+ all layers} \\
LLaMA 3-paradigm transfer & N/A & \textbf{MW$\rightarrow$IC AUC $0.805$ (L25)} \\
LLaMA 3-paradigm factor decomp & 2-paradigm: 69.5\% & \textbf{3-paradigm: 75.8\%} \\
Cross-bet-type transfer (F1) & Not tested & \textbf{LLaMA: AUC $0.736$--$0.927$ (all $p=0.000$)} \\
Cross-bet-type transfer (F1) & Not tested & \textbf{Gemma SAE: AUC $0.696$--$0.902$ (deep layers)} \\
Common BK features (bet-type) & Not quantified & \textbf{415 features (213 promoting, 202 inhibiting)} \\
LLaMA 2D BK subspace & Not computed & \textbf{IC AUC$=0.901$, SM AUC$=0.943$} \\
LLaMA G-prompt alignment & Behavioral only ($1.26\times$) & \textbf{$\cos(G_\mathrm{dir}, BK_\mathrm{dir})$: L22$=+0.634$, L30$=+0.548$} \\
LLaMA bet constraint mapping & Not computed & \textbf{$r=0.987$, $p=0.013$} \\
\bottomrule
\end{tabular}
\end{table}

\end{document}
